\documentclass[../tsr2102.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}
\fancyhf[LH]{How To Use This Book}
\begin{document}
\noindent{The new \textit{Monstrous Compendium} format was designed with the
Dungeon Master in mind. Monster entries are given in alphabetical order,
printed on looseleaf sheets that can be organized for convenience. The
sheets needed for an adventure can be placed in a separate folder for play,
then later returned to the master binder. New monsters can be easily
placed in the master binder.}\\
\indent{All monsters described here are typical for their type; likewise, the
given encounter tables are guidelines for general play. DMs should note
that unusual variations are encouraged, but are most effective when they
depart from the expected.}\\
\indent{Each monster is described fully, with entries that describe behavior,
combat modes, and so on. These are explained below.}\\

\noindent{CLIMATE/TERRAIN defines where the creature is most often found. Climates include arctic, subarctic, temperate, and tropical. Typical terrain
includes plain/scrub, forest, rough/hill, mountain, swamp, and desert.}\\
FREQUENCY is the likelihood of encountering a creature in an area. Very
rare is a 4\% chance, rare is 11\%, uncommon is 20\%, and common is a
65\% chance. Chances can be adjusted for special areas.
ORGANIZATION is the general social structure the monster adopts.
“Solitary” includes small family groups.
ACTIVE CYCLE is the time of day when the monster is most active. Those
most active at night may be active at any time in subterranean settings.
Active cycle is a general guide and exceptions are fairly common.
DIET shows what the creature generally eats. Carnivores eat meat, herbivores eat plants, and omnivores eat either. Scavengers eat mainly cartion.
INTELLIGENCE is the equivalent of human “IQ”. Certain monsters are
instinctively cunning; these are noted in the monster descriptions. Ratings
correspond roughly to the following Intelligence ability scores:
0 Non-intelligent or not ratable
A Animal intelligence
2-4 Semi-intelligent
5-7 _ Low intelligence
8-10 Average (human) intelligence
11-12 ‘Very intelligent
13-14 Highly intelligent
15-16 Exceptionally intelligent
17-18 Genius
19-20 Supra-genius
21+ — Godlike intelligence
TREASURE refers to the treasure tables in the Dungeon Masters Guide. If
individual treasure is indicated, each individual may carry it (or not, at
the DM's discretion). Major treasures are usually found in the monster's
lair; these are most often designed and placed by the DM. Intelligent monsters will use magical items present and try to carry off their most valuable
treasures if hard pressed. If treasure is assigned randomly, roll for each
type possible: if all rolls fail, no treasure of any type is found. Treasure
should be adjusted downward if few monsters are encountered. Large
treasures are noted by a parenthetical multiplier (x 10, etc.)—not to be
confused with treasure type X. Do not use the tables to place dungeon
treasure, as numbers encountered underground will be much smaller.
ALIGNMENT shows the general behavior of the average monster of that
type. Exceptions, though uncommon, may be encountered.
NO. APPEARING indicates an average encounter size for a wilderness
encounter, The DM should alter this to fit the circumstances as the need
arises. This should not be used for dungeon encounters.
ARMOR CLASS is the general protection worn by humans and humanoids, protection due to physical structure or magical nature, or difficulty
in hitting due to speed, re! , etc. Humans and humanoids of roughly
man-size that wear armor will have an unarmored rating in parentheses.
Listed AC do not include any special bonuses noted in the description.
MOVEMENT shows the relative speed rating of the creature. Higher
speeds may be possible for short periods. Human, demihuman, and humanoid movement rate is often determined by armor type (unarmored
rates are given in parentheses). Movements in different mediums are abbreviated as follows: Fl = fly, Sw = swim, Br = burrowing, Wb = web.
Flying creatures will also have a Maneuverability Class from A to E.
HIT DICE controls the number of hit points damage a creature can withstand before being killed. Unless otherwise stated, Hit Dice are 8-sided
(1-8 hit points). The Hit Dice are rolled and the numbers shown are added
to determine the monster's hit points. Some monsters will have a hit point
spread instead of Hit Dice, and some will have additional points added to
their Hit Dice. Thus, a creature with 4 +4 Hit Dice has 4d8 +4 hit points
(8-36 total). Note that creatures with +3 or more hit points are considered
the next higher Hit Die for purposes of attack rolls and saving throws.
THACO is the attack roll the monster needs to hit armor class 0. This is always a function of Hit Dice, except in the case of very large, nonaggressive herbivores (such as some dinosaurs). Humans and
demihumans always use player character THACOs, regardless of whether
they are player characters or “monsters”. THACOs do not include any
special bonuses noted in the descriptions.
NUMBER OF ATTACKS shows the basic attacks the monster can make in
a melee round, excluding special attacks. This number can be modified by
hits that sever members, spells such as haste and slow, and so forth. Multiple attacks indicate several members, raking paws, multiple heads, etc.
DAMAGE PER ATTACK shows the amount of damage a given attack will
make, expressed as a spread of hit points (dice roll combination). If the
monster uses weapons, the damage will done by the typical weapon will
be followed by the parenthetical note “weapon”. Damage bonuses due to
Strength are listed as a bonus following the damage range.
SPECIAL ATTACKS detail attack modes such as dragon breath, magic
use, etc. These are explained in the monster description.
SPECIAL DEFENSES are precisely that, and are detailed in the monster
description.
MAGIC RESISTANCE is the percentage chance that magic cast upon the
creature will fail to affect it, even if other creatures nearby are affected. If
the magic penetrates the resistance, the creature is still entitled to any normal saving throw allowed.
SIZE is abbreviated as: “T,” tiny (2' tall or less); “S,” smaller than a typical
human (2 +’ to 4’); “M,” man-sized (4+' to 7’); “L,” larger than man-sized
(7+! to 12'); “Hy” huge (12+! to 25’); and “G,” gargantuan (25 +’).
MORALE is a general rating of how likely the monster is to persevere in
the face of adversity or armed opposition. This guideline can be adjusted
for individual circumstances. Morale ratings correspond to the following
2-20 range:
2-4 Unreliable
5-7 Unsteady
8-10 Average
11-12 Steady
13-14 Elite
15-16 Champion
17-18 Fanatic
19-20 Fearless
XP VALUE is the number of experience points awarded for defeating (not
neccessarily killing) the monster. This value is a guideline that can be
modified by the DM for the degree of challenge, encounter situation, and
for overall campaign balance.
Combat is the part of the description that discusses special combat abilities, arms and armor, and tactics.
Habitat/Society outlines the monster's general behavior, nature, social
structure, and goals.
Ecology describes how the monster fits into the campaign world, gives
any useful products or byproducts, and presents other miscellaneous information.
Close variations of a monster (e.g. merrow, ogre) are given in a special
section after the main monster entry. These can be found by consulting the
index to find the major listing.
\end{document}